%-------------------------------------------------------------------------------
% yum_references
%-------------------------------------------------------------------------------
%
% \file        yum_references.tex
% \library     Documents
% \author      Chris Ahlstrom
% \date        2015-06-02
% \update      2017-09-24
% \version     $Revision$
% \license     $XPC_GPL_LICENSE$
%
%     Provides the References section of yoshimi-user-manual.tex.  Rather
%     than use the bibtex package, our small set of references uses a
%     simpler method.
%
%-------------------------------------------------------------------------------

\section{References}
\label{sec:yumreferences}

   This section provides references for this \textsl{Yoshimi} user's manual, as
   well as some other information.
   
   Although the \textsl{Yoshimi} project is based on SourceForge, it also has a
   mirror on GitHub, and a mailing list on FreeLists.Org.
   One can subscribe with an e-mail to:
   \url{yoshimi-request@freelists.org}
   or by visiting: \url{http://www.freelists.org/list/yoshimi}.
   To post to the list, email to: \url{yoshimi@freelists.org}
   The archive of the old SourceForge mailing list is found
   at: \url{https://www.freelists.org/archive/yoshimi}.

% \subsection{Web Sites}
% \label{subsec:references_web_sites}
%
% \iffalse
%  \begin{enumber}
%     \item \href{http://yoshimi.sourceforge.net/}{Yoshimi Download Site}
%        \url{http://yoshimi.sourceforge.net/}
%     \item \href{http://rakkarrak/}{Rakkarrak Guitar Effects}
%  \end{enumber}
% \fi

% If we wrap the URLs in \url{}, then we get web-color errors on CentOS!

\begin{thebibliography}{99}

   \bibitem{bankrootupgrades}
   Will J. Godfrey
   \emph{A discussion of making Bank/Root specifications more regular.}
   \url{http://sourceforge.net/p/yoshimi/mailman/message/33200765/}
   2014.

   \bibitem{book}
   Chris Ahlstrom
   \emph{A Yoshimi User's Manual}
   \url{https://github.com/ahlstromcj/yoshimi-doc/}
   2015.

   \bibitem{cookbook}
   Chris Ahlstrom
   \emph{A Yoshimi Cookbook}
   \url{https://github.com/ahlstromcj/yoshimi-cookbook/}
   2015.

   \bibitem{cookbookeq}
   Robert Bristow-Johnson.
   \emph{Cookbook formula for audio EQ biquad filter coefficients}
   \url{http://www.musicdsp.org/files/Audio-EQ-Cookbook.txt}
   2005.  The "files" directory has a number interesting files as wel..

   \bibitem{cormi}
   Cormi
   \emph{Cormi Collection}
   \url{https://www.freesound.org/people/cormi/}
   His cormi57.wordpress.com site has a lot of other nice sound material, as
   well.
   2015.

   \bibitem{drumsds}
   Straulino
   \emph{A collection of instruments}
   \url{http://www.straulino.ch/zynaddsubfx/}
   \url{http://www.straulino.ch/zynaddsubfx/Drums\_DS\_v2012-12-08.zip}
   \url{http://www.straulino.ch/zynaddsubfx/ZynAddSubFX\_Natural\_Drum\_Kit\_Demo\_v2012-06-22.ogg}

   \bibitem{folderol}
   "folderol"
   \emph{A collection of instruments}
   \url{http://www.kara-moon.com/forum/index.php?topic=762.0}

   \bibitem{godfrey}
   Will Godfrey
   \emph{Will Godfrey's Music}
   \url{http://www.musically.me.uk}
   See the "Bits 'n Stuff" link especially.

   \bibitem{gpl2vsgpl3}
   Will J. Godfrey
   \emph{A discussion of licensing issues with Youshimi and ZynAddSubFX
   components, and FLTK library versions.}
   \url{http://sourceforge.net/p/yoshimi/mailman/yoshimi-devel/}
   2015.

   \bibitem{kvraudio}
   caonoize
   \emph{ZynAddSubFX by paulnasca: Downloads, Banks, Patches, etc.}
   \url{http://www.kvraudio.com/product/zynaddsubfx\_by\_paulnasca/downloads}
   2015.

   \bibitem{mmxgn}
   mmxgn
   \emph{Instruments made with zynaddsubfx/yoshimi}
   \url{https://github.com/mmxgn/instruments-zyn}
   2011.

   \bibitem{rakarrack}
   Rakarrack team
   \emph{The download site for the Rakarrack software effects engine.}
   \url{http://rakarrack.sourceforge.net/}
   2015.

   \bibitem{ringmodulator}
   LinuxMusicians newsgroup
   \emph{Ring Modulation in ZynAddSubFX}
   \url{http://linuxmusicians.com/viewtopic.php?f=1&t=8178}
   2012.

   \bibitem{scala}
   Manuel Op de Coul <\url{coul@huygens-fokker.org}>
   \emph{The Scala Musical Tuning Application.}
   \url{http://www.huygens-fokker.org/scala/}
   Scala is a powerful software tool for experimentation with musical
   tunings, such as just intonation scales, equal and historical
   temperaments, microtonal and macrotonal scales, and non-Western scales.
   2014.

   \bibitem{sharphall}
   Sharphall
   \emph{How to create drum sounds in ZynAddSubFX or Yoshimi, Part 1}
   \url{http://sharphall.org/docs/zynaddsubfx\_yoshimi\_drum\_tutorial.php}
   Never got continued, unfortunately.

   \bibitem{synthsecrets}
   Gordon Reid
   \emph{Synth Secrets:  Creative Synthesis}
   \url{http://www.soundonsound.com/sos/allsynthsecrets.htm}
   1999-2004.

   \bibitem{x31eq}
   Graham
   \emph{A small musical website}
   \url{http://x31eq.com} and specifically
   \url{http://x31eq.com/zasf}

   \bibitem{yoshimi}
   Yoshimi team \url{abrolag@users.sourceforge.net}
   \emph{The download site for the Yoshimi software synthesizer.}
   \url{http://yoshimi.sourceforge.net/}
   2015.

   \bibitem{yoshimi2}
   Yoshimi team
   \emph{The alternate location for the Yoshimi source-code.}
   \url{https://github.com/abrolag/yoshimi now.}
   2015.

   \bibitem{yoshiminews}
   Yoshimi team.
   \emph{Yoshimi user/developer mailing list}
   \url{http://www.freelists.org/list/yoshimi}
   2015.

   \bibitem{yoshiminewsarchive}
   Yoshimi team.
   \emph{Yoshimi user/developer mailing list archive}
   \url{http://www.freelists.org/archive/yoshimi}
   2015.

   \bibitem{zynaddsubfx}
   Mark McCurry, Paul Nasca (ZynAddSubFX team)
   \emph{The download site for the ZynAddSubFX software synthesizer.}
   \url{http://zynaddsubfx.sourceforge.net/}
   2015.

   \bibitem{zynbanks2009}
   Paul Nasca?
   \emph{The download site for the ZynAddSubFX banks, instruments,
      parameters, and demos}
   \url{http://zynaddsubfx.sourceforge.net/doc/instruments/}
   2009.

   \bibitem{zyncmake}
   ZynAddSubFX team.
   \emph{Building ZynAddSubFX}
   \url{http://zynaddsubfx.sourceforge.net/Doc/#_appendix_b_building_zynaddsubfx}
   2009.

   \bibitem{zyndemos}
   AMSynth team.
   \emph{Provides a number of OGG demos of ZynAddSubFX sounds. It
   also includes the author's own "Vanilla" bank, and links to additional
   patch collections and demonstration videos.}
   \url{http://www.amsynth.com/zynaddsubfx.html}

   \bibitem{zyndoc}
   Mark McCurry, Paul Nasca
   \emph{ZynAddSubFX online manual.}
   \url{http://zynaddsubfx.sourceforge.net/Doc}
   2015.

   \bibitem{zynodt}
   Paul Nasca
   \emph{Original ZynAddSubFX manual, ODT format.}
   \url{http://linux.autostatic.com/docs/zynaddsubfx\_manual-v0.1.odt}
   2011.

   \bibitem{zynpdf}
   Paul Nasca
   \emph{Original ZynAddSubFX manual, PDF format.}
   \url{http://linux.autostatic.com/docs/zynaddsubfx\_manual-v0.1.pdf}
   2011.

   \bibitem{zynsource}
   Paul Nasca, Mark Murray
   \emph{The download package for the ZynAddSubFX source-code}
   \url{http://downloads.sourceforge.net/project/zynaddsubfx/zynaddsubfx/2.5.0/zynaddsubfx-2.5.0.tar.gz}

   \bibitem{zynwiki}
   Jeremy ("autostatic")
   \emph{ZynAddSubFX Manual, good overall of most Zyn settings and knobs}
   \url{http://wiki.linuxaudio.org/wiki/zynaddsubfx_manual}
   2011.

\end{thebibliography}

%-------------------------------------------------------------------------------
% vim: ts=3 sw=3 et ft=tex
%-------------------------------------------------------------------------------
