%-------------------------------------------------------------------------------
% yum_command_line_v2
%-------------------------------------------------------------------------------
%
% \file        yum_command_line_v2.tex
% \library     Documents
% \author      Chris Ahlstrom
% \date        2015-09-06
% \update      2018-09-23
% \version     $Revision$
% \license     $XPC_GPL_LICENSE$
%
%     Provides the description of the no-gui mode of Yoshimi.
%
%-------------------------------------------------------------------------------

\section{The Yoshimi Command-Line Interface}
\label{sec:command_line}

   \index{CLI}
   \index{cmd}
   \index{command line}
   \textsl{Yoshimi} provides a command-line mode of operation
   where many aspects of the application can be controlled via text commands.
   This mode is useful for blind people and for programmers, for example.
   These text commands can be put into a script file, and that script can then
   be run.
   
   To access the command-line mode,
   add the \texttt{-i} or \texttt{--no-gui} command-line
   option when starting \textsl{Yoshimi} on the command-line.  But note that,
   when starting \textsl{Yoshimi} on the command-line without the "no gui"
   option, the "command-line" mode
   of operation is available at the same time as the GUI, as well.

   One of the main features of recent \textsl{Yoshimi} releases
   is improved non-GUI accessibility.  In fact,
   \textsl{Yoshimi} can run with neither GUI nor CLI input access. Working
   purely as a hidden MIDI device, a daemon of sorts. To enable a tidy close,
   there is a new short-form NRPN. Just send 68 to both MSB and LSB (CC 99 and
   CC98).

   In a command line environment, almost all the 'running' commands are
   available, and now most of the instrument editing ones are available.
   The whole of vector control, and much of MIDI-learn, is also exposed to the
   command line.

   One can decide what MIDI/audio setup is wanted, list and set roots and
   banks, load instruments into any part, change a part's channel, set main
   volume and key shift, and set up vector control.  A number of first-time
   defaults have been changed to make this feature easier.

   When starting from the command line, an argument can be included for a new
   root path to be defined to point to a set of banks fetched from elsewhere.
   This will be given the next free ID. A future upgrade will allow the ID to
   be set to any valid one when it is created, mirroring the GUI behaviour.

   Once running, almost all configuration can be done within the terminal
   window.  There is also extensive control of roots, banks, parts and
   instruments including the ability to list and set all of these.

   Additional controls that are frequently taken for granted in the GUI, but
   otherwise get forgotten, are \textsl{master key shift} and \textsl{master
   volume}.

%  The whole of vector control, and much of MIDI-learn, is also exposed to the
%  command line.

   The command-line mode provides extensive error checking and feedback.
   Note the change in nomenclature from "Parameters" to "Patch Set", which is
   visible in the main screen, and also reflected in the command line.
   \index{command level}
   The prompt will always show what \textsl{command level}
   one is on, along with relevant information.
   At the CLI prompt, when effects are being managed, the preset number is also
   shown on the prompt so you'll typically see something like:

   \begin{verbatim}
      yoshimi part 2 FX 1 Rever-7 >
   \end{verbatim}

   One will also get a confirmation message, but, for clarity,
   those are not included in the examples below.
   Here is an example session:

   Starting from the \texttt{yoshimi} prompt:

   \begin{verbatim}
      yoshimi> s p 2 pr 107
      yoshimi part 2 >
   \end{verbatim}

   This command \textbf{s}ets \textbf{p}art number 2 to \textbf{pr}ogram
   number 107 from the \textsl{current} instrument bank.
   \textsl{Yoshimi} is now on part 2 as the current part (indicated by the
   prompt), and all subsequent commands will relate to this "level".
   At this level, one can change the current part simply with:

   \begin{verbatim}
      yoshimi part 2 > s 4
      yoshimi part 4 >
   \end{verbatim}

   \textsl{Yoshimi} is now on part number 4. Now set an effect:

   \begin{verbatim}
      yoshimi part 4 > s ef ty re
      yoshimi part 4 FX 1 rever-1 >
   \end{verbatim}

   This command \textbf{s}ets the part's \textbf{ef}fect 1 (implicit) to
   \textbf{ty}pe \textbf{re}verb.

   Note that many settings parameters are optional, and if you omit them,
   either a default or last-used value will be assumed. Also, names are
   truncated to 5 characters so the prompt line doesn't get unmanageably long.

   From here you can set a preset for this effect:

   \begin{verbatim}
      yoshimi part 4 FX 1 rever > s pre 3
   \end{verbatim}

   Since V 1.5.0 the \textbf{pre}sets have been shown in the prompt, and one
   will still get a confirmation message.

   Settings that follow in a direct command "path" through several levels
   can be made all at once, and one will be left at the appropriate level.
   Thus, summarizing some of the above commands:

   \begin{verbatim}
      s p 4 ef 2 ty re
      yoshimi part 4 FX 2 rever >
   \end{verbatim}

   One cannot combine \texttt{type} and \texttt{preset} as they
   are both at the same level.  To go back one level, use the
   ".." command (reminiscent of the \texttt{cd ..} operation in an OS command
   shell):

   \begin{verbatim}
      yoshimi part 4 FX 2 rever > ..
      yoshimi part 4 >
   \end{verbatim}

   To go back to the top command level, use the "/" command:

   \begin{verbatim}
      yoshimi part 4 > /
      yoshimi >
   \end{verbatim}

   These two special level-movement commands can also be put on the front of
   any other command.  Starting where we were before:

   \begin{verbatim}
      yoshimi part 4 FX 2 rever > .. s vol 70
      yoshimi part 4 >
   \end{verbatim}

   Part 4 volume is now at 70, and \textsl{Yoshimi} is once again at the
   "part level", not the "part FX level".
%  Also note that the space after the ".." is optional.

   The help menus and lists are also partially context sensitive. This
   feature should help avoid clutter and confusion.

   As well as an immediate history, \textsl{Yoshimi} maintains a single command
   history file for all instances of \textsl{Yoshimi} that records any
   non-duplicated loads or save.  Thus, provided one makes a normal
   command-line exit, the last commands will be available on the next run of
   \textsl{Yoshimi}.
   The command-line now has formal methods of opening, selecting and closing
   additional instances.

   When loading external files from the command line, there is an alternative
   to entering the full name if \textsl{Yoshimi} has already seen this file and
   it is in the history list. In this situation you enter '@' followed by the
   list number.

   \begin{verbatim}
      yoshimi> l h v

      Recent Vectors:
      1  /home/will/another.xvy
      2  /home/will/Subtle.xvy
      3  /home/will/excellent.xvy
      4  /home/will/yoshimi-code/examples/CoolSound.xvy

      yoshimi> lo ve @4
      Main Vector Loaded /home/will/yoshimi-code/examples/CoolSound.xvy to chan 1
   \end{verbatim}

   The loading of externally-saved instruments is also done, by default,
   relative to one's \textsl{Yoshimi} home directory.  However, saving an
   external instrument from the command-line still requires a full pathname.

   \index{cmd!at-sign}
   The 'recent history' lists can now load MIDI-learned files, patchsets, or
   vector files numerically from the associated list instead of having to type
   the names out.  This uses the '@' (list number) operator.

   Originally this section described the currently implemented commands,
   but as the command set is very much a moving target, it is simpler to just
   ask one to run \textsl{Yoshimi} and type the "?" command.

   Commands with "*" in the description need the setup to be saved,
   and \textsl{Yoshimi} restarted to be activated.

   Note that \textsl{Yoshimi}'s command-line can also load and save states,
   patchsets, and scales, and can list recent histories.
   Vector load and save is also supported from the command-line.
   That's a complete vector set, with all mappings, instruments, etc.
   One can independently decide which
   channel to load and save from, so that one
   can actually build up a vector set in
   (say) channel 3, then later decide to use it in channel 7.
   It has the extension \texttt{.xvy}, standing for "Xml/Vector/Yoshimi".
   It will eventually be integrated with the saved states.

   Another small detail is that all of the minimum command-line
   abbreviations are now \textbf{C}apitalised in the help list.

   More features will be added, and the organisation of them may be changed
   slightly.  If any configuration settings are changed, either at the
   command-line or in the graphical user-interface, one will be given a warning
   when exiting, with the option to continue running so one can save the
   changes.

\subsection{Command Level}
\label{subsec:command_line_command_level}

   \index{cli!command level}
   \index{cli!context level}
   A command level (also known as a "context level")
   is simply a position in the hierarchy of commands that cover
   some aspect of \textsl{Yoshimi} functionality.
   The levels are:

   \begin{itemize}
      \item \textbf{Top Level}
      \item \textbf{System Effects}
      \item \textbf{Insertion Effects}
      \item \textbf{Part}
      \item \textbf{Part Effects}
      \item \textbf{Scales (microtonal)}
      \item \textbf{Vector}
      \item \textbf{Controllers}
      \item \textbf{Synth Engines}
      \begin{itemize}
         \item \textbf{Addsynth}
         \item \textbf{Addsynth Voice}
         \item \textbf{Subsynth}
         \item \textbf{Padsynth Harmonics}
         \item \textbf{Padsynth Envelopes}
      \end{itemize}
   \end{itemize}

   Any level that has direct numerical content can be changed with
   "set (n)" once at that level.  The level is
   indicated by the text in the \textsl{Yoshimi} prompt.
   For example, one can set 1 to 16 vector channels, so from the
   \textsl{Top} level, the following command
   will set the default  (1, or the last-used number):

   \begin{verbatim}
      set vector                          # or "s ve"
   \end{verbatim}

   Given this level (the \textsl{Vector} level), the following command
   will switch to vector channel 5.

   \begin{verbatim}
      set 5                               # or "s 5"
   \end{verbatim}

   However, at the start, one could have gone straight to 5 with:

   \begin{verbatim}
      set vector 5                        # or "s ve 5"
   \end{verbatim}

   A detailed discussion of command-line vector control is presented in
   \sectionref{subsection:vector_command_line}.

\subsection{Command Scripts}
\label{subsec:command_line_command_scripts}

   \textsl{Yoshimi} command-line users can run plain-text scripts that behave
   in exactly the same way as if the commands had been entered from the
   command-line directly.
   The actual script command can be initiated from any context/level
   and is simply:

   \begin{verbatim}
      RUN {filepath-of-script}
   \end{verbatim}

   To avoid confusion, the script routine first sets the context to the
   top level, then performs all the commands, and returns to the original
   level. If there is a fault in the script, it will be reported along with the
   number of the line where the error occurred.
   Here is a simple example:

   \begin{verbatim}
      set part on
      set add voice on
      set volume 45
   \end{verbatim}

   This script makes sure the part is on, that the relative addsynth voice
   is on, and finally sets its volume to 45.

   Although this process starts from the top level, it will use the parameters
   that were last set.
   Thus, if one had been working on part 7, addsynth voice 2,
   then that is the one that will have its volume adjusted.
   This means one can set up generic preferences,
   then apply them to any part, engine, etc.

   The script routine honors any normal abbreviations.
   Blank lines are ignored.
   There is no comment facility; this feature may be added later.

\subsection{Command Depth}
\label{subsec:command_line_command_depth}

   Recent developments in \textsl{Yoshimi} have made it possible to greatly
   (one could say dramatically) extend command-line access deep into
   the synth structures.
   This creates a problem where the command line itself could become
   unmanageably long.
   Thus, now only the current context level is printed in full.
   The levels higher up the tree are minimized:

   \begin{verbatim}
      yoshimi Part 1+>
      yoshimi P1, Sub>
      yoshimi P1, S, analog Filter>
   \end{verbatim}

   Rather than stating that a switch is on or off,
   there is now just a \texttt{+} sign for "on", and nothing for "off".
   This is clearer than using a \textsl{-}, and the
   slight shift in the line gives the user another visual clue.

   There is a new command in the "config" context that controls where this is
   displayed, or whether it is shown at all.
   This command is:
   
   \begin{verbatim}
      EXPose {OFf, ON, PRompt}
   \end{verbatim}

   \textbf{Off} will give the bare prompt with no other information.
   \textbf{On} shows it as a separate line above the prompt:

   \begin{verbatim}
           @ P1, S, analog Filter
           yoshimi>
   \end{verbatim}

   \textbf{Prompt} shows it as a part of the prompt:

   \begin{verbatim}
           yoshimi P1, S, analog Filter>
   \end{verbatim}

   The default setting is \textbf{ON}.

\subsection{Commands for Synth Engines and Parts}
\label{subsec:command_line_command_synth_engines}

   \textsl{Yoshimi} has a number of commands for controlling and
   configuring the the synth engines from the command-line.

   For the new commands, first of all there is the part kit structure.
   There are three modes that the kits (i.e. settings)
   of the engines can take:
   
   \begin{enumerate}
      \item Multi
      \item Single
      \item Crossfade
   \end{enumerate}

   These forms are exactly the same as the graphical controls,
   and can be set once in the part context.
   Starting at the part level prompt, this command will return the setting on
   the line after the command,
   and show a new context level prompt:

   \begin{verbatim}
      yoshimi Part 1+> set kmode multi
      Part 1 Kit Mode multi
      yoshimi Part 1+, kit 1+, multi>
   \end{verbatim}

   This setting is at the kit item 1 (which is always enabled).
   Let's change to kit item 4, and, since
   kit item 4 hasn't yet been enabled, let's enable it and turn on the
   SubSynth engine so that it will sound:

   \begin{verbatim}
      yoshimi Part 1+, kit 1+, multi> set kitem 4
      yoshimi Part 1+, kit 4, multi>
      yoshimi Part 1+, kit 4, multi> set on
      yoshimi Part 1+, kit 4+, multi> set sub on
      yoshimi P1+, K4+, M, Sub+>
   \end{verbatim}

   Note how the prompt line is more compact, and indicates via the
   plus-signs that the part, kit, and subsynth are all on.

   These new controls are shown in the part context help list.
   The synth engines, AddSynth, Voice, SubSynth and PadSynth also have their
   own contexts with appropriate help lists.
   There is also a new help list which doesn't show directly in any context,
   but has to be called explicitly.
   It is:
   
   \begin{verbatim}
      ? COMmon
   \end{verbatim}
   
   and lists all the commands that are
   applicable from the part level through the new engine contexts.
   A typical example is "Volume" which is a context-sensitive control for part,
   AddSynth, Voice, SubSynth, and PadSynth.
   LFO, Filter, and Envelope have their own contexts above
   whichever engine they are sitting on, and again have their own help lists.
   Currently the greatest context depth is:

   \begin{verbatim}
      yoshimi P1+, K4+, M, A+, V5+, freq LFO>
   \end{verbatim}

   Envelopes and LFOs are slightly complicated as they have groupings, and one
   must specify the command first, followed by the group, but they are
   displayed the other way round:

   \begin{verbatim}
      yoshimi P1+, K4+, M, A+, Voice 5+> set lfo frequency
   \end{verbatim}

   This command would set the context shown above, and the need for
   compression is obvious.
   At that level, the \texttt{?} command will list the LFO controls.

\subsubsection{Part Common Commands}
\label{subsec:command_line_part_common_commands}

   \begin{table}[H]
      \centering
      \caption{Part Common Commands}
      \label{table:yoshimi_part_common_commands}
      \begin{tabular}{l l}

\texttt{ON @} &
   Enables the part/kit item/engine/insert etc. \\
\texttt{OFf @} &
   Disables as above.   \\
\texttt{Volume n @} &
   Volume.  \\
\texttt{Pan n2 @} &
   Panning. \\
\texttt{VElocity n @} &
   Velocity sensing sensitivity. \\
\texttt{MIn n +} &
   Minimum MIDI note value.   \\
\texttt{MAx n +} &
   Maximum MIDI note value.   \\
\texttt{DETune Fine n *} &
   Fine frequency.   \\
\texttt{DETune Coarse n *} &
   Coarse stepped frequency.  \\
\texttt{DETune Type n *} &
   Type of coarse stepping.   \\
\texttt{OCTave n *} &
   Shift octaves up or down.  \\
\texttt{FIXed s *-add} &
   Set base frequency to 440Hz (ON, other).  \\
\texttt{EQUal n *-add} &
   Equal temper variation. \\
\texttt{BENd Adjust n  *-add} &
   Pitch bend range. \\
\texttt{BENd Offset n  *-add} &
   Pitch bend shift. \\
\texttt{STEreo s *-voice} &
   ON, (other).   \\
\texttt{DEPop n \&} &
   Initial attack slope.   \\
\texttt{PUnch Power n} \& &
   Attack boost amplitude. \\
\texttt{PUnch Duration n \&} &
   Attack boost time.   \\
\texttt{PUnch Stretch n \&} &
   Attack boost extend. \\
\texttt{PUnch Velocity n \&} &
   Attack boost velocity sensitivity.  \\
\texttt{OVertone Position s \#} &
   Relationship to fundamental, see below. \\
\texttt{OVertone First n \#} &
   Degree of first parameter. \\
\texttt{OVertone Second n \#} &
   Degree of second parameter.   \\
\texttt{OVertone Harmonic n \#} &
   Amount harmonics are forced.  \\
\texttt{LFO ... *-sub} &
   Enter LFO insert context.  \\
\texttt{FILter ... *} &
   Enter Filter insert context.  \\
\texttt{ENVelope ... *} &
   Enter Envelope insert context.   \\

      \end{tabular}
   \end{table}

Notes:

   \begin{enumerate}
      \item \texttt{@} Exists in all part contexts.
      \item \texttt{+} Part and kit mode controls.
      \item \texttt{*} Add, Sub, Pad and AddVoice controls.
      \item \texttt{*-add} Not AddSynth.
      \item \texttt{*-sub} Not SubSynth.
      \item \texttt{*-voice} Not AddVoice.
      \item \texttt{\&} AddSynth \& PadSynth only.
      \item \texttt{\#} SubSynth \& PadSynth only.
   \end{enumerate}

Overtone Position s, Relationship to fundamental, values:

   \begin{itemize}
      \item HArmonic
      \item SIne
      \item POwer
      \item SHift
      \item UShift
      \item LShift
      \item UPower
      \item LPower
   \end{itemize}

\subsubsection{Part Commands}
\label{subsec:command_line_part_commands}

   Note that n1 is the part number in the table that follows.

   \begin{table}[H]
      \centering
      \caption{Part Commands}
      \label{table:yoshimi_part_commands}
      \begin{tabular}{l l}

\texttt{OFfset n2} &
   Sense offset.  \\
\texttt{Breath s} &
   Control (ON, {other}).  \\
\texttt{POrtamento s} &
   (ON, other).   \\
\texttt{Mode s} &
   Mode (Poly, Mono, Legato). \\
\texttt{Note n2} &
   Polyphony.  \\
\texttt{SHift n2} &
   Shift semitones (0 no shift). \\
\texttt{EFfects [n2]} &
   Effects context level.  \\
 \texttt{(effect) Type s} &
   Effect type.   \\
 \texttt{(effect) PREset n3} &
   Numbered effect preset to n3. \\
 \texttt{(effect) Send n3 n4} &
   Part to system effect n3 at volume n4. \\
\texttt{KMode s} &
   Part to kit mode (MUlti, SIngle, CRoss, OFf).   \\
 \texttt{(kmode) KItem n} &
   Kit item number (1-16). \\
   \texttt{(kitem) MUte s} &
      This item (ON, {other}).   \\
   \texttt{(kitem) KEffect n} &
      Effect for this item (0-none, 1-3). \\
 \texttt{DRum s} &
   Kit to drum mode (ON, {other}).  \\
\texttt{PRogram n2/s} &
   Instrument ID / CLear sets default. \\
\texttt{NAme s} &
   The display name the part can be saved with. \\
\texttt{Channel n2} &
   Channel ( 32 disables,  16 note off only).   \\
\texttt{Destination s2} &
   Audio destination (Main, Part, Both).  \\
\texttt{ADDsynth ...} &
   Enter AddSynth context. \\
\texttt{SUBsynth ...} &
   Enter SubSynth context. \\
\texttt{PADsynth ...} &
   Enter PadSynth context. \\
\texttt{? COMmon} &
   Controls common to most part contexts. \\

      \end{tabular}
   \end{table}

   The following commands enter context for the various types of
   synth engines:

   \begin{itemize}
      \item \textbf{Part AddSynth}:
        VOice ... Enter Addsynth voice context.
      \item \textbf{Part AddVoice}:
        WAveform ... Enter the oscillator waveform context.
      \item \textbf{Part PadSynth}:
      \begin{itemize}
         \item APply. Puts the latest changes into the wavetable.
         \item WAveform ... Enter the oscillator waveform context.
      \end{itemize}
   \end{itemize}

\paragraph{Part SubSynth Commands}
\label{paragraph:command_line_part_subsynth_commands}

   \begin{table}[H]
      \centering
      \caption{Part SubSynth Commands}
      \label{table:yoshimi_part_subsynth_commands}
      \begin{tabular}{l l}

\texttt{HArmonic n1 Amp n2} &
   Set harmonic n1 to n2 intensity. \\
\texttt{HArmonic n1 Band n2} &
   Set harmonic n1 to n2 width. \\
\texttt{HArmonic Stages n} &
   Number of stages. \\
\texttt{HArmonic Mag n} &
   Harmonics filtering type. \\
\texttt{HArmonic Position n} &
   Start position. \\
\texttt{BAnd Width n} &
   Common bandwidth. \\
\texttt{BAnd Scale n} &
   Bandwidth slope v frequency. \\

      \end{tabular}
   \end{table}

\paragraph{Part Waveform Commands}
\label{paragraph:command_line_part_waveform_commands}

   \begin{table}[H]
      \centering
      \caption{Part Waveform Commands}
      \label{table:yoshimi_part_waveform_commands}
      \begin{tabular}{l l}

   \texttt{HArmonic n1 Amp n2} &
      Set harmonic n1 to n2 intensity. \\
   \texttt{HArmonic n1 Phase n2} &
      Set harmonic n1 to n2 phase. \\
   \texttt{CLear} &
      Clear harmonic settings. \\
   \texttt{SHape s} &
      Set the shape of the basic waveform. \\
   \texttt{APply} &
      Fix settings (only for PadSynth). \\

      \end{tabular}
   \end{table}

   This list shows the "shape of the basic waveform" settings available.

   \begin{itemize}
      \item SIne
      \item TRiangle
      \item PUlse
      \item SAw
      \item POwer
      \item GAuss
      \item DIode
      \item ABsine
      \item PSine
      \item SSine
      \item CHIrp
      \item ASine
      \item CHEbyshev
      \item SQuare
      \item SPike
      \item Circle
   \end{itemize}

\subsubsection{Engine Envelopes}
\label{subsec:command_line_engine_envelopes}

   \begin{table}[H]
      \centering
      \caption{Engine Envelopes, Type}
      \label{table:yoshimi_engine_envelopes_types}
      \begin{tabular}{l l}

\texttt{AMplitude} &
   Amplitude type. \\
\texttt{FRequency} &
   Frequency type. \\
\texttt{FIlter} &
   Filter type. \\
\texttt{BAndwidth} &
   Bandwidth type (SubSynth only). \\

      \end{tabular}
   \end{table}

   \begin{table}[H]
      \centering
      \caption{Engine Envelopes, Controls}
      \label{table:yoshimi_engine_envelopes_controls}
      \begin{tabular}{l l}

\texttt{Expand n} &
   Overall envelope time. \\
\texttt{Force s} &
   Force release (ON, other). \\
\texttt{Linear s} &
   Linear slopes (ON, other). \\
\texttt{FMode s} &
   Set as Freemode (ON, other). \\

      \end{tabular}
   \end{table}

   \begin{table}[H]
      \centering
      \caption{Engine Envelopes, Fixed}
      \label{table:yoshimi_engine_envelopes_fixed}
      \begin{tabular}{l l}

\texttt{Attack Level n} &
   Initial attack level. \\
\texttt{Attack Time n} &
   Time before decay point. \\
\texttt{Decay Level n} &
   Initial decay level. \\
\texttt{Decay Time n} &
   Time before sustain point. \\
\texttt{Sustain n} &
   Sustain level. \\
\texttt{Release Time n} &
   Time to actual release. \\
\texttt{Release Level n} &
   Level at envelope end. \\

      \end{tabular}
   \end{table}

Example: "S FR D T 40" is "set frequency decay time 40".

Note: Some envelopes have limited controls.

   \begin{table}[H]
      \centering
      \caption{Engine Envelopes, Freemode}
      \label{table:yoshimi_engine_envelopes_freemode}
      \begin{tabular}{l l}

\texttt{Points} &
   Number of defined points (read only). \\
\texttt{Sustain n} &
   Point number where sustain starts. \\
\texttt{Insert n1 n2 n3} &
   Insert point at 'n1' with X increment 'n1', Y value 'n2'. \\
\texttt{Delete n} &
   Remove point 'n'. \\
\texttt{Change n1 n2 n3} &
   Change point 'n1' to X increment 'n1', Y value 'n2'. \\

      \end{tabular}
   \end{table}

\subsubsection{Engine Filters}
\label{subsec:command_line_engine_filters}

   \begin{table}[H]
      \centering
      \caption{Engine Filters}
      \label{table:yoshimi_engine_filters}
      \begin{tabular}{l l}

\texttt{CEnter n} &
   Center frequency. \\
\texttt{Q n} &
   Q factor.   \\
\texttt{Velocity n} &
   Velocity sensitivity.   \\
\texttt{SLope n} &
   Velocity curve.   \\
\texttt{Gain n} &
   Overall amplitude.   \\
\texttt{TRacking n} &
   Frequency tracking.  \\
\texttt{Range s} &
   Extended tracking (ON, {other}). \\
\texttt{CAtegory s} &
   Analog, Formant, State variable. \\
\texttt{STages n} &
   filter stages (1 to 5). \\
\texttt{TYpe s} &
   Category dependent, and not formant filters.  See types below.   \\

      \end{tabular}
   \end{table}

   The list of filter types:

   \begin{itemize}
      \item Analog filters:
      \begin{itemize}
          \item \textbf{l1}. One stage low pass.
          \item \textbf{h1}. One stage high pass.
          \item \textbf{l2}. Two stage low pass.
          \item \textbf{h2}. Two stage high pass.
          \item \textbf{band}. Two stage band pass.
          \item \textbf{stop}. Two stage band stop.
          \item \textbf{peak}. Two stage peak.
          \item \textbf{lshelf}. Two stage low shelf.
          \item \textbf{hshelf}. Two stage high shelf.
      \end{itemize}
      \item State variable filters:
      \begin{itemize}
         \item \textbf{low}. Low pass.
         \item \textbf{high}. High pass.
         \item \textbf{band}. Band pass.
         \item \textbf{stop}. Band stop.
      \end{itemize}
   \end{itemize}

   \begin{table}[H]
      \centering
      \caption{Engine Filters, Formant Editor}
      \label{table:yoshimi_engine_filters_formant_editor}
      \begin{tabular}{l l}

\texttt{Direct n} &
   Bypass LFOs, envelopes. \\
\texttt{FCenter n} &
   Center frequency of sequence. \\
\texttt{FRange n} &
   Octave range of formants. \\
\texttt{Expand n} &
   Stretch overal sequence time. \\
\texttt{Lucidity n} &
   Clarity of vowels. \\
\texttt{Morph n} &
   Speed of change between formants. \\
\texttt{SIze n} &
   Number of vowels in sequence. \\
\texttt{COunt n} &
   Number of formants in vowels. \\
\texttt{VOwel n} &
   Vowel being processed. \\
\texttt{Point n1 n2} &
   Vowel n1 at sequence position n2. \\

      \end{tabular}
   \end{table}

\subsubsection{Engine LFOs}
\label{subsec:command_line_engine_lfos}

   \begin{table}[H]
      \centering
      \caption{Engine LFOs}
      \label{table:yoshimi_engine_lfos}
      \begin{tabular}{l l}

\texttt{AMplitude} &
   Amplitude type. \\
\texttt{FRequency} &
   Frequency type. \\
\texttt{FIlter} &
   Filter type. \\
\texttt{Rate n} &
   Frequency. \\
\texttt{Start n} &
   Start position in cycle. \\
\texttt{Delay n} &
   Time before effect. \\
\texttt{Expand n} &
   Overall LFO time. \\
\texttt{Continuous s} &
   ON, other. \\
\texttt{Type s} &
   LFO oscillator shape. \\
\texttt{AR n} &
   Amplitude randomness. \\
\texttt{FR n} &
   Frequency randomness. \\

      \end{tabular}
   \end{table}

  e.g. S FI T RU              set filter type ramp up

Filter types (s parameter, LFO oscillator shape):

   \begin{itemize}
      \item SIne
      \item Triangle
      \item SQuare
      \item RUp (ramp up)
      \item RDown (ramp down)
      \item E1dn
      \item E2dn
   \end{itemize}

\subsection{Other Command Tables}
\label{subsec:command_line_other_command_tables}

   When running from the command line, commands
%  (see \tableref{table:yoshimi_text_commands_part_1}
%  and \tableref{table:yoshimi_text_commands_part_2})
   can be entered after the 'up and running' message.
   Commands are \textsl{not} case-sensitive.
   Commands can be abbreviated to the first three letters of each command,
   or, in some cases, just one letter.  This is indicated by
   uppercase letters in command descriptions.
   The commands available depend on the current "context" of the command line.
   However, there is a group of commands always available:

   \begin{itemize}
      \item \texttt{\textbf{?}} or \texttt{\textbf{help}}
      \item \texttt{\textbf{L}ist}
      \item \texttt{\textbf{RES}et}
      \item \texttt{\textbf{EX}it}
   \end{itemize}

   Apart from these commands, the command line works on a system of
   context levels, and normally only the commands relevant to that "level"
   will be available.

%  This whole section needs a rewrite unfortunately, as it has changed
%  dramatically. The good news is that the new format has the seal of approval
%  from blind users, and so is very unlikely to change. However, new list and
%  context levels will be added once Will starts implementing the part engine
%  controls in a near-future version.

   We describe the command lists here. These lists are relative to
   a particular context, and what one sees if one enters
   \textbf{?} while at that level, to get help.
   However the command lists can all be called \textsl{specifically} from any
   level. From any higher level, \textbf{? ?} will show the top level one.

\subsubsection{Top Commands}
\label{subsec:command_line_top_command_list}

   These commands are part of the Top context/command level.
   First, one gets the default options, always available.
   Then there are several options that have ellipsis (...);
   these are the context submenus.
   After that come all the actual top level controls; there are still a lot!

   \begin{verbatim}
      yoshimi> ?
   \end{verbatim}

   Note that there are a number of commands common to all command levels.
   We describe them here.

%  \begin{table}[H]
%     \centering
%     \caption{Yoshimi Common Commands}
%     \label{table:yoshimi_text_common_commands}
%     \begin{tabular}{l l}

\begin{center}
% \begin{longtable}{l l}
\begin{longtable}{p{2cm} p{10cm}}
\caption[Yoshimi Common Commands]{Yoshimi Common Commands} \\

\texttt{?,  Help} &
   Show commands.  Also note specialized versions such as \textbf{? list}. \\
\texttt{STop} &
   All sound off.  Stop!  Panic! \\
\texttt{RESet} &
   Return to start-up conditions (if answering 'y' to the prompt).
   The \textbf{RESet} command asks for confirmation, and, if allowed, will
   reset \textsl{Yoshimi} to startup conditions.
   However, this will \textsl{not} clear any MIDI-learned lines. \\
\texttt{EXit} &
   Tidy up and close Yoshimi (if 'y' to the prompt). \\
\texttt{..} &
   Step up one command level. \\
\texttt{/} &
   Step up to the top command level. \\

\end{longtable}
\end{center}

%     \end{tabular}
%  \end{table}

%  The \textbf{RESet} command asks for confirmation, and, if allowed, will
%  reset \textsl{Yoshimi} to startup conditions.
%  However, note that this will \textsl{not} clear any MIDI-learned lines.
%  The \textbf{EXit} command will ask for confirmation, and, if allowed, will
%  perform a clean exit.

   \begin{table}[H]
      \centering
      \caption{Yoshimi Top-Level Commands}
      \label{table:yoshimi_text_toplevel_commands}
      \begin{tabular}{l l}

\texttt{Part [n1] ...} &
   Start part operations.  See below. \\
\texttt{VEctor [n1] ...} &
   Start vector operations. \\
\texttt{SCale ...} &
   Scale (microtonal) operations. \\
\texttt{LIst ...} &
   Show various available parameters. \\
\texttt{LOad ...} &
   Load various files. \\
\texttt{SAve ...} &
   Save various files. \\
\texttt{COnfig ...} &
   Engage configuration settings. \\
\texttt{ADD} &
   Add paths and files. \\
\texttt{ADD Root [s]} &
   Add a root path [s] to the root list. \\
\texttt{ADD Bank [s]} &
   Add a bank [s] to the current root. \\
\texttt{ADD YOshimi [n]} &
   Start a new instance. \\
\texttt{IMport [s [n1]] [n2] [s]} &
   Imports a bank. \\
\texttt{EXport [s [n1]] [n2] [s]} &
   Exports a bank. \\
\texttt{REMove} &
   Remove paths, files, and entries. \\
\texttt{REMove Root [n]} &
   De-list root path ID [n]. \\
\texttt{REMove Bank [n]} &
   Delete bank ID [n] (and all its contents) from the current root. \\
\texttt{REMove YOshimi [n]} &
   Close an instance. \\
\texttt{REMove MLearn [s] [n]} &
   Delete MIDI learned values: 'ALL' removes whole list, or select line [n]. \\
\texttt{Set / Read} &
   Set or read all main parameters. See below for context levels. \\
\texttt{Set/read Root [n]} &
   Read the current root path or set it to ID [n]. \\
\texttt{Set/read Bank [n]} &
   Read the current bank or set the current bank to ID [n]. \\
\texttt{Set/read YOshimi [n]}&
   Switch to instance [n] for further control. \\
\texttt{Set/read MLearn [n] [s]} &
   MIDI learned line number [n] control.  See below. \\
\texttt{SYStem effects [n]} &
   System effects for editing. \\
\texttt{SYStem effects SEnd [n2] [n3]} &
   Send system effect to effect [n2] at volume [n3]. \\
\texttt{SYStem effects preset [n2]} &
   Set effect preset to number [n2]. \\
\texttt{INSert effects [n1]} &
   Add an nsertion effect [n1] for editing. \\
\texttt{INSert effect SEnd [s]/[n2]} &
   Set effect [s] where [n2] = (Master, Off or part number). \\
\texttt{Insert effect PREset [n2]} &
   Set numbered effect preset to [n2]. \\
\texttt{AVailable [n]} &
   Set the available number of parts, [n] = 16, 32, 64. \\
\texttt{Volume [n]} &
   Set Master volume to [n]. \\
\texttt{SHift [n]} &
   Master key shift in semitones [n] (0 no shift). \\
\texttt{DEtune [n]} &
   Set the master fine detune to [n] to match other sound sources. \\
\texttt{SOlo [n]} &
   Set the channel 'solo' switcher t [n]
   (0 = off, 1 = row, 2 = col, 3 = loop) \\
\texttt{SCC [n]} &
   Set incoming 'solo' channel number to [n]. \\
\texttt{TIMes [s]} &
   Time display on instrument load message (ENable / other). \\
\texttt{'...'} &
   Help sub-menu. \\

      \end{tabular}
   \end{table}

\paragraph{SOlo}
\label{paragraph:command_line_solo}

   The \textbf{SOlo [s] [n]} and \textbf{SCC [n]}
   commands enable and set \textsl{Yoshimi}'s 'Solo' feature,
   whereby one can silently switch MIDI input to different parts.
   The [n] parameter can be 0 = off, 1 = row, 2 = col, and 3 = loop).
   'Row' and 'Loop' mode use the first 16 parts, while
   'Column' mode can use all possibly 64 parts.

   This setting has to be decided before setting 'CC', which then determines
   which channel to listen to for performing the actual switch.

   See \sectionref{subsec:mixer_panel_window}; it goes into more details about
   this setting, at a user-interface level.

\paragraph{Set / Read Context Levels}
\label{paragraph:command_line_context_levels}

   The Set / Read commands set or read all main parameters.  There are a few
   commands that set the context or command level, where additional commands
   peculiar to the "context" are provided.  Here are the command/context levels
   (also see \sectionref{subsec:command_line_command_level}.)
   Note that we also list commands for the effects levels.

   \begin{itemize}
      \item \texttt{Part}. Enter context level for part operations.
      \item \texttt{VEctor}. Enter context level for vector operations.
      \item \texttt{SCale}. Enter context level for scale (microtonal) operations.
      \item \texttt{MLearn}. Enter context level for MIDI Learn operations.
      \item \texttt{COnfig}. Enter context level for configuration settings.
      \item \texttt{SYStem effects [n]}. Enter effects context level.
      \begin{itemize}
         \item \texttt{Type [s]}.  Set the effect type.
         \item \texttt{PREset [n2]}. Set the numbered effect preset to n2.
         \item \texttt{SEnd [n2] [n3]}. Send a system effect to effect n2
            at volume n3.
      \end{itemize}
      \begin{itemize}
   \item \texttt{INSert effects [n1]}. Enter effects context level.
         \item \texttt{Type [s]}. Set the effect type.
         \item \texttt{PREset [n2]}. Set numbered effect preset to n2.
         \item \texttt{SEnd [s]/[n2]}. Set where to send the effect
            ('Master', 'Off', or a part number).
      \end{itemize}
   \end{itemize}

\paragraph{Part Command Level}
\label{paragraph:command_line_context_level_part}

   This command switches to the part context level and makes all its commands
   accessible.  If no number '[n]' is entered it will be on the default part
   (1) or whatever was the previous part in use.
   See \sectionref{subsec:command_line_part_command_list}, which goes into
   details.

\paragraph{MLearn}
\label{paragraph:command_line_mlearn}

   The commands for MLearn are the following values for the [s] parameter:

   \begin{itemize}
      \item \texttt{MUte}, which Enables or, for any other provide token
         disables this line.
      \item \texttt{7Bit} will Enable incoming (learned) NRPNs as a 7-bit value.
      \item \texttt{CC [n2]} will set the incoming controller value that will
         be recognized.  This command can re-order the list.
      \item \texttt{CHan [n2]} will set the incoming channel number that will
         be recognized.  This command can re-order the list.
      \item \texttt{MIn [n2]} will set the conversion for the incoming
         value to a minimum percentage.
      \item \texttt{MAx [n2]} will set the conversion for the incoming
         value to a maximum percentage.
      \item \texttt{LImit [s]} set to Enable will use limiting instead of
         compression.  The conversion uses the minimum and maximum limits.
      \item \texttt{BLock [s]} set to Enable inhibits other lines on this
         CC/channel pair.  It prevents this CC/channel pair from being passed
         on to any other lines or system controls.  It has not effect if the
         line has been disabled.
   \end{itemize}

\subsubsection{Part Commands}
\label{subsec:command_line_part_command_list}

   Next we have the part list. It's now much clearer than it was.
   Notice it was called from the top level with '? part':

   \begin{verbatim}
      yoshimi> ? part
   \end{verbatim}

   In the table that follow, we leave off the following commands, already noted
   above in the first table
   (see \sectionref{subsec:command_line_top_command_list}):
   \textbf{?}, \textbf{Help}, \textbf{STop}, \textbf{RESet},
   \textbf{EXit}, \textbf{..}, and \textbf{/}.

   Note that, for a part, [n1] is the part number.

   \begin{table}[H]
      \centering
      \caption{Yoshimi Part Commands}
      \label{table:yoshimi_text_part_commands}
      \begin{tabular}{l l}

\texttt{ENable} &
   Enables the part and MIDI and virtual keyboard access. \\
\texttt{DIsable} &
   Disables the part. \\
\texttt{Volume [n2]} &
   Set the part's volume to [n2]. \\
\texttt{Pan [n2]} &
   Set part's L/R panning to [n2]. 64 = center. \\
\texttt{VElocity [n2]} &
   Set part velocity sensing sensitivity to [n2]. \\
\texttt{OFfset [n2]} &
   Set velocity sense offset (start point) to [n2]. \\
\texttt{POrtamento [s]} &
   Set portamento. ([s] = Enable, other values disable it). \\
\texttt{Mode [s]} &
   Set keying mode ([s] = Poly, Mono, Legato). \\
\texttt{Note [n2]} &
   Set maximum note polyphony value to [n2]. \\
\texttt{SHift [n2]} &
   Set key shift semitones to [n2] (0 = no shift). \\
\texttt{MIn [n2]} &
   Set the minimum MIDI note value accepted to [n2]. \\
\texttt{MAx [n2]} &
   Set the maximum MIDI note value accepted to [n2]. \\
\texttt{EFfects [n2]} &
   Enter Effects command level, and set effects number to [n2], if given. \\
\texttt{Effects Type [s]} &
   Set the effect type to [s]. \\
\texttt{Effects PREset [n3]} &
   Set the numbered effect preset to [n3]. \\
\texttt{Effects Send [n3] [n4]} &
   Send part to system effect [n3] at volume [n4]. \\
\texttt{PRogram [n2]} &
   Loads instrument ID [n2] in the current bank. \\
\texttt{NAme [s]} &
   Sets the display name the part can be saved with to [s]. \\
\texttt{Channel [n2]} &
   Set part's MIDI channel ([n2] > 32 disables, [n2] > 16 note-off only). \\
\texttt{Destination [s2]>} &
   JACK audio destination ([s2] = Main, Part, Both) \\

      \end{tabular}
   \end{table}

   For the \textbf{Part Name} command, this sets the display name the part can
   be saved with, either in a bank or externally.
   For the \textbf{Part Channel} command, this sets the incoming MIDI channel
   the part recognises. If it is set greater than 32, then all messages will be
   ignored.  If it is set greater than 16, then only Note-Off messages will be
   recognised.
   For the \textbf{Part Destination} command, for JACK audio only,
   this sets where this part's audio will be sent. It can be the main output
   pair, the part's output pair or both.

\subsubsection{Vector Commands}
\label{subsec:command_line_vector_command_list}

   The vector list is called from the part level with '? ve'.

   \begin{verbatim}
      yoshimi part 1 on> ? ve
   \end{verbatim}

   The commands at this level deal with control of an X axis and a Y axis.
   The CC for the X axis must be set before everything else.  Then the CC for
   the Y axis must be set.  Finally, the other Y controls can be set.

   In the table that follows, we leave off the commands already noted
   above in the first table
   (see \sectionref{subsec:command_line_top_command_list}).

   Note that for vector, [n1] = base channel.

   \begin{table}[H]
      \centering
      \caption{Yoshimi Vector Commands}
      \label{table:yoshimi_text_vector_commands}
      \begin{tabular}{l l}

\texttt{VEctor [n1]} &
   Vector channel operations. \\
\texttt{[X/Y] CC [n2]} &
   CC [n2] is used for theX or Y axis sweep. \\
\texttt{[X/Y] Features [n] [s]} &
   Sets X or Y features [n] = 1 to 4 ([s] = Enable, Reverse, other = off).  \\
\texttt{[X] PRogram [l/r] [n2]} &
   Sets X program change ID [n2] for left or right part. \\
\texttt{[Y] PRogram [d/u] [n2]} &
   Sets Y program change ID [n2] forL DOWN or UP part. \\
\texttt{[X/Y] Control [n2] [n3]} &
   Sets [n3] CC to use for X or Y feature [n2] = 2 to 4. \\
\texttt{Off} &
   Disables vector control for this channel.  Parts are unchanged. \\
\texttt{Name [s]} &
   Sets the internal name for this complete vector.  \\

      \end{tabular}
   \end{table}

   The \textbf{X/Y Features [n2]} command sets the features for the
   selected axis, and also if they are to be off or reversed.
   See the \textbf{Features} section,
   \sectionref{subsec:command_line_command_descriptions}, for more information.
   It seems to indicate a feature range of 1 to 4, not 2 to 4.

   The \textbf{X/Y Control [n2] [n3]} command
   sets the [n3] CC to use for the X or Y feature [n2] = 2 to 4.
   This allows a change of the actual CC associated with features 2 through 4.
   They can be any CC that \textsl{Yoshimi} recognises.

\subsubsection{Scales Commands}
\label{subsec:command_line_scales_command_list}

   A completely new one is the scales list:

   \begin{verbatim}
      yoshimi> ? sc
   \end{verbatim}

   In the table that follow, we leave off the following commands, already noted
   above in the first table
   (see \sectionref{subsec:command_line_top_command_list}):
   \textbf{?}, \textbf{Help}, \textbf{STop}, \textbf{RESet},
   \textbf{EXit}, \textbf{..}, and \textbf{/}.

   \begin{table}[H]
      \centering
      \caption{Yoshimi Scales Commands}
      \label{table:yoshimi_text_scales_commands}
      \begin{tabular}{l l}

\texttt{FRequency [n]} &
   Set the 'A' note's actual frequency to [n], usually 440 Hz. \\
\texttt{NOte [n]} &
   Set the 'A' note's number to [n]. \\
\texttt{Invert [s]} &
   Invert the entire scale ([s] = enable, other values = off) \\
\texttt{CEnter [n]} &
   Set the note number of the key's center to [n]. \\
\texttt{SHift [n]} &
   Shift the entire scale up or down by [n]. \\
\texttt{SCale [s]} &
   Activate microtonal scale ([s] = enable, other values = off). \\
\texttt{MApping [s]} &
   Activate keyboard mapping ([s] = enable, other values = off). \\
\texttt{FIrst [n]} &
   Set the first note number to be mapped to [n]. \\
\texttt{MIddle [n]} &
   Set the middle note number to be mapped to [n]. \\
\texttt{Last [n]} &
   Set the last note number to be mapped to [n]. \\
\texttt{Tuning [s] [s2]} &
   Set the CSV tuning values (see below). \\
\texttt{Keymap [s] [s2]} &
   Set the CSV keymap (n1, n2, n3, etc.); s2 = 'IMPort' from named file. \\
\texttt{NAme [s]} &
   Set the internal name for this scale. \\
\texttt{DEscription [s]} &
   Sets the description of this scale. \\
\texttt{CLEar} &
   Clear all settings and revert to the standard scale. \\

      \end{tabular}
   \end{table}

\paragraph{Scales / Tuning [s] [s2]}
\label{paragraph:command_line_tuning}

   Tuning sets the CSV tuning values, which are decimal numbers or ratios
   (n1.n1 or n1/n1, n2.n2 or n2/n2, etc.).
   The [s2] parameter requests an 'IMPort' from a named file, which is
   \index{.scl}
   \index{config!.scl}
   a \texttt{.scl} file.

\paragraph{Scales / Keymap [s] [s2]}
\label{paragraph:command_line_keymap}

   Keymap either sets the keyboard mapping values as a comma separated list, or
   imports a
   \index{.kbm}
   \index{config!.kbm}
   \texttt{.kbm} file from a named file [s2].

\subsubsection{Help List}
\label{subsec:command_line_help_list}

   You can now clearly see which items can be listed with:

   \begin{verbatim}
      yoshimi> ? li
      yoshimi> ? list
      yoshimi> help
   \end{verbatim}

   In the table that follow, we leave off commands noted above
   (see \sectionref{subsec:command_line_top_command_list}).

   \begin{table}[H]
      \centering
      \caption{Yoshimi Help Commands}
      \label{table:yoshimi_text_help_commands}
      \begin{tabular}{l l}

\texttt{Roots} &
   List all available root paths. \\
\texttt{Banks [n]} &
   List the banks in root ID [n] or the current root.  See below. \\
\texttt{Instruments [n]} &
   List instruments in bank ID [n] or current bank. See below.  \\
\texttt{Parts} &
   List parts with instruments installed. \\
\texttt{Vectors} &
   List settings for all enabled vectors. \\
\texttt{Settings} &
   List dynamic settings. \\
\texttt{Tuning} &
   Microtonal scale tunings. See the Scales section. \\
\texttt{Keymap} &
   Microtonal scale keyboard map.  See the Scales section. \\
\texttt{Config} &
   Show the current configuration. See the Config section. \\
\texttt{MLearn [s[n]]} &
   MIDI learned controls ('@n' for full details on one line). \\
\texttt{History [s]} &
   Show recent files. See below. \\
\texttt{Effects [s]} &
   Show effect types ([s] = 'all' includes preset numbers and names). \\
\texttt{PREsets} &
   Show all the presets for the currently selected effect. \\

      \end{tabular}
   \end{table}

A few more detailed descriptions occur in the following sections.

\paragraph{List / Banks [n]}
\label{paragraph:command_line_list_banks}

   This command shows all of the banks present in either the numbered ([n])
   bank root, or in the current one (if no number is provided).

\paragraph{List / Instruments [n]}
\label{paragraph:command_line_list_instruments}

   This command shows all of the instruments present in either the numbered
   ([n]) bank root, or in the current one (if no number is provided).

\paragraph{List / History [s]}
\label{paragraph:command_line_list_history}

   Show the recent history of the following items [s]:
   Instruments, Patchsets, SCales, STates, Vectors, and MLearn).
   If no parameter is provided, show them all.

   The last-used file in any section is now always at the top of its history list,
   so it's easier to pick up where one left off.
   Instruments, patch sets, vectors, scales, MIDI-learn and state all offer the
   most recent entry whenever one wants to load or save.  On first-time use (when
   there is no history) the home directory will be offered as a location,
   regardless of where \textsl{Yoshimi} was called from.

   In the specific case of instruments, when \textsl{saving},
   one is offered the instrument in the currently-selected part to the home
   directory, otherwise, when saving these 'managed' files,
   one won't be offered the previous last-used entry unless it was seen on that
   session, either by being loaded, or saved by name. This is to give some
   protection against accidental overwrites.

   For example:
   You have been working on the 'foo' instrument
   for a whole day, saving the whole patch set as you
   go. Then, the following day, you start up \textsl{Yoshimi}
   and immediately have a completely
   new idea 'bar' and start working on it. Without thinking, you save and hit
   Enter. Oops, your just wiped out 'foo'.
   Only now you haven't, because while loading \textsl{Yoshimi}
   would see the older file, so that saving
   will offers your home directory to put a new name in.

\paragraph{List / Effects [s]}
\label{paragraph:command_line_list_effects}

   Show the effect types.
   [s] = 'All' includes preset numbers and names.
   If this command is called from the Effects level, then one see only the name
   of the current effect and the number of presets.

\subsubsection{Load/Save List}
\label{subsec:command_line_loadsave_list}

   And the same for load and save:

   \begin{verbatim}
      yoshimi> ? lo
      yoshimi> ? sa
   \end{verbatim}

   In the table that follow, we leave off the commands noted above
   (see \sectionref{subsec:command_line_top_command_list}).

   \begin{table}[H]
      \centering
      \caption{Yoshimi Load Commands}
      \label{table:yoshimi_text_load_commands}
      \begin{tabular}{l l}

\texttt{Instrument [s]} &
   Load instrument to current part from a named file [s]. \\
\texttt{SCale [s]} &
   Load and activate scale settings from named file [s]. \\
\texttt{VEctor [n] [s]} &
   Load and activate vector to channel n (or saved) from named file [s]. \\
\texttt{Patchset [s]} &
   Load and activate a complete patch set from named file [s]. \\
\texttt{MLearn [s]} &
   Load the full MIDI learned list from named file [s]. \\
\texttt{STate [s]} &
   Load all system settings and patch sets from named file [s]. \\

      \end{tabular}
   \end{table}

   For the \textbf{Load Instrument} command, the instrument is enabled if it is
   configured to be enabled.  For the \textbf{Load Vector} command, if there is
   no number parameter, the vector is loaded to the channel it was originally
   saved from.
   For the \textbf{Load Patchset} command, all instruments, scales, and vectors
   are loaded from the named file.
   For the \textbf{Load STate} command, all configuration, system settings,
   patch sets, and MIDI-learned lines are loaded from the named file.
   These notes also apply to the \textbf{Save} version of these commands.

   \begin{table}[H]
      \centering
      \caption{Yoshimi Save Commands}
      \label{table:yoshimi_text_loadsave_commands}
      \begin{tabular}{l l}

\texttt{Instrument [s]} &
   Save current part to named file [s]. \\
\texttt{SCale [s]} &
   Save current scale settings to named file [s]. \\
\texttt{VEctor [n] [s]} &
   Save vector on channel n to named file [s]. \\
\texttt{Patchset [s]} &
   Save complete set of instruments to named file [s]. \\
\texttt{MLearn [s]} &
   Save midi learned list to named file [s]. \\
\texttt{STate [s]} &
   Save all system settings etc. to named file [s].  See above. \\
\texttt{Config} &
   Save current configuration. \\

      \end{tabular}
   \end{table}

\subsubsection{Config Commands}
\label{subsec:command_line_config_list}

   Finally there is the new and shiny (and quite big) COnfig command level:

   \begin{verbatim}
      yoshimi> ? con
   \end{verbatim}

   In the table that follow, we leave off the commands noted above
   (see \sectionref{subsec:command_line_top_command_list}).  Also
   note that more complete descriptions follow this table.

   \begin{table}[H]
      \centering
      \caption{Yoshimi Config Commands}
      \label{table:yoshimi_text_config_commands}
      \begin{tabular}{l l}

\texttt{Oscillator [n]} &
   * Add/Pad size (power 2 256-16384). \\
\texttt{BUffer [n]} &
   * Internal size (power 2 16-4096). \\
\texttt{PAdsynth [s]} &
   Interpolation type (Linear, other = cubic). \\
\texttt{Virtual [n]} &
   Keyboard layout (0 = QWERTY, 1 = Dvorak, 2 = QWERTZ, 3 = AZERTY). \\
\texttt{Xml [n]} &
   Set the XML compression level to [n] (0-9). \\
\texttt{REports [s]} &
   Destination for reporting (Stdout, other = console). \\
\texttt{STate [s]} &
   * Autoload default at start (Enable; other = disable). \\
\texttt{Hide [s]} &
   Hide non-fatal errors (Enable; other = disable). \\
\texttt{Display [s]} &
   GUI splash screen (Enable; other = disable). \\
\texttt{Time [s]} &
   Add to instrument load message (Enable; other = disable). \\
\texttt{Include [s]} &
   Include XML headers on file load (Enable; other = disable). \\
\texttt{Keep [s]} &
   Include inactive data on all file saves (Enable; other = disable). \\
\texttt{Gui [s]} &
   * Run with GUI (Enable, Disable). \\
\texttt{Cli [s]} &
   * Run with CLI (Enable, Disable). \\
\texttt{MIdi [s]} &
   * Connection type (Jack, Alsa). \\
\texttt{AUdio [s]} &
   * Connection type (Jack, Alsa). \\
\texttt{ALsa Midi [s]} &
   * Name of ALSA MIDI source. \\
\texttt{ALsa Audio [s]} &
   * Name of ALSA audio hardware device. \\
\texttt{ALsa Sample [n]} &
   * ALSA sampling rate (0 = 192000, 1 = 96000, 2 = 48000, 3 = 44100). \\
\texttt{Jack Midi [s]} &
   * Name of JACK MIDI source. \\
\texttt{Jack Server [s]} &
   * Name of JACK server. \\
\texttt{Jack Auto [s]} &
   * Connect JACK on start (Enable; other = disable). \\
\texttt{ROot [n]} &
   Root CC (0 - 119, other disables). \\
\texttt{BAnk [n]} &
   Bank CC (0, 32, other disables). \\
\texttt{PRogram [s]} &
   MIDI program change enabling (0 is disable, other is enable). \\
\texttt{ACtivate [s]} &
   MIDI program change activates part (0 is off, other is on). \\
\texttt{EXtended [s]} &
   Extended program change (0 is off, other is enable). \\
\texttt{Quiet [s]} &
   Ignore 'reset all controllers' (Enable other). \\
\texttt{Log [s]} &
   Log incoming MIDI CCs (Enable other). \\
\texttt{SHow [s]} &
   GUI MIDI learn editor (Enable other). \\

      \end{tabular}
   \end{table}

   '*' marks entries that need to be saved, and \textsl{Yoshimi}
   restarted, to activate them.

\paragraph{Config / Oscillator [n]}
\label{paragraph:command_line_config_oscillator}

   This sets the size of the buffer used for both AddSynth and PadSynth
   oscillators, and is always a power of 2.

\paragraph{Config / BUffer [n]}
\label{paragraph:command_line_config_buffer}

   This is the size of the audio buffer that \textsl{Yoshimi} uses.  For ALSA
   audio, it will always be the same size as ALSA's buffering, but for JACK it
   can be the same, bigger, or smaller. It also is always a power of 2.

\paragraph{Config / PAdsynth [n]}
\label{paragraph:command_line_config_padsynth}

   Sets the quality of the intepolation that PadSynth uses on its wavetables.
   'Linear' is faster, but 'Cubic' is (potentially) very slightly better
   quality.

\paragraph{Config / Virtual [n]}
\label{paragraph:command_line_config_virtual}

   This setting controls the layout of the virtual keyboard, and can match the
   commonest computer keyboards.

\paragraph{Config / Xml [n]}
\label{paragraph:command_line_config_xml}

   Sets the amount of compression used on all
   \textsl{Yoshimi}'s data files. 9 is the most-compressed setting.
   0 is no compression, so that the configuration file
   can be read in an ordinary text editor.

\paragraph{Config / REports [s]}
\label{paragraph:command_line_config_reports}

   Determines where almost all information and error messages will be sent. A
   few will always go to \texttt{stderr} (such as the ones reporting a GUI
   problem).

\paragraph{Config / STate [s]}
\label{paragraph:command_line_config_state}

   Sets whether a pre-saved default state file will be loaded on start-up.

\paragraph{Config / Hide [n]}
\label{paragraph:command_line_config_hide}

   Sets to ignore non-fatal system errors, or verbose messages.

\paragraph{Config / Display [s]}
\label{paragraph:command_line_config_display}

   Enables \textsl{Yoshimi}'s start-up splash screen (which is
   enabled at first time start).

\paragraph{Config / Time [s]}
\label{paragraph:command_line_config_time}

   Attaches the time an instrument took to load and initialize to the loading
   message.

\paragraph{Config / Include [s]}
\label{paragraph:command_line_config_include}

   Reports version information on all data failures.

\paragraph{Config / Keep [s]}
\label{paragraph:command_line_config_keep}

   Sets up to innclude all data on file saves, including data for
   inactive and random elements.

\paragraph{Config / Gui [s]}
\label{paragraph:command_line_config_gui}

   Run with the graphic user interface enabled.

\paragraph{Config / Cli [s]}
\label{paragraph:command_line_config_cli}

   Run with the command-line interface enabled.

\paragraph{Config / MIdi [s]}
\label{paragraph:command_line_config_midi}

   Sets whether MIDI input comes from JACK or from ALSA.

\paragraph{Config / AUdio [s]}
\label{paragraph:command_line_config_audio}

   Sets whether audio is passed out to JACK or ALSA.

\paragraph{Config / ALsa Midi [s]}
\label{paragraph:command_line_config_alsa_midi}

   Sets the name of an ALSA MIDI source to which
   \textsl{Yoshim}i will try to connect.

\paragraph{Config / ALsa Audio [s]}
\label{paragraph:command_line_config_alsa_audio}

   Sets the name of a hardware audio device to which ALSA will try to connect.

\paragraph{Config / ALsa Sample [n]}
\label{paragraph:command_line_config_alsa_sample}

   Sets the sampling rate when using ALSA audio.

\paragraph{Config / Jack Midi [s]}
\label{paragraph:command_line_config_jack_midi}

   Sets the name of a JACK MIDI source to which
   \textsl{Yoshimi} will try to connect.

\paragraph{Config / Jack Server [s]}
\label{paragraph:command_line_config_jack_server}

   Sets the name of an audio server to which JACK will try to connect.

\paragraph{Config / Jack Auto [s]}
\label{paragraph:command_line_config_jack_auto}

   Determines whether JACK will try to connect the main L=R audio outputs at
   start-up time.

\paragraph{Config / ROot [n]}
\label{paragraph:command_line_config_root}

   Provides the MIDI CC that \textsl{Yoshimi} expects bank root changes to come
   from.

\paragraph{Config / BAnk [n]}
\label{paragraph:command_line_config_bank}

   Provides the MIDI CC that \textsl{Yoshimi} expects bank changes to come from.

\paragraph{Config / PRogram [s]}
\label{paragraph:command_line_config_program}

   Determines whether MIDI program changes are honored or ignored.

\paragraph{Config / ACtivate [s]}
\label{paragraph:command_line_config_activate}

   Enables a part when it gets a MIDI program change message,
   if it was disabled.

\paragraph{Config / Extended [s]}
\label{paragraph:command_line_config_extended}

   Sets a MIDI CC for receiving program changes for the top (extra)
   32 instruments in a bank.

\paragraph{Config / Quiet [s]}
\label{paragraph:command_line_config_quiet}

   Sets up to ignore a MIDI 'reset all controllers' message.

\paragraph{Config / Log [s]}
\label{paragraph:command_line_config_log}

   Displays the value of received MIDI CCs.

\paragraph{Config / SHow [s]}
\label{paragraph:command_line_config_show}

   A setting for the GUI MIDI learn editor, where [s] is 'Enable', or some
   other token to disable the feature.
   This setting indicates to automatically open the MIDI-learn editor window
   when a sucessful 'learn' has been made.

\subsection{Command Descriptions}
\label{subsec:command_line_command_descriptions}

   This section describes the command-line commands in more detail.
   Obviously, some more needs to be written about some of the commands.
   Note that all the parameters for the \texttt{load} and \texttt{save}
   parameters are strings, and the parameters are compulsory, not optional.

   \setcounter{ItemCounter}{0}      % Reset the ItemCounter for this list.

   \itempar{..}{cmd!up one level}
      \index{cmd!..}
      Step back up one command context level.
      This command can immediately precede another command, so that the second
      command takes places at the context above the current context.

   \itempar{/}{cmd!to top level}
      \index{cmd!/}
      Step back up to the top command context level.
      This command can immediately precede another command, so that the second
      command takes places at the top context.

   \itempar{add bank [s]}{cmd!add bank}
      Define a new bank, \textsl{s}, where \textsl{s} is a bank name,
      and add it to the current root.

   \itempar{remove bank [s]}{cmd!remove bank}
      Delete the bank named \textsl{s}, and all its contents,
      from the current root path.

   \itempar{export bank [s [n1]] [n2] [s]}{cmd!export bank}
      The command line now has two commands to provide access to the new bank
      export and import controls. These are top level controls and are used as
      below.  The command above is used to export a bank. The square bracket term
      is optional, and enables one to select a different root to export from and
      would be in the form:

\begin{verbatim}
   EXport Root (root ID number) (bank ID number) (full path name to export to)
\end{verbatim}

   If one is happy to export from the currently selected root, then this
   simplifes to:

\begin{verbatim}
   EXport (bank ID number) (full path name to export to)
\end{verbatim}

   \itempar{import bank [s [n1]] [n2] [s]}{cmd!import bank}
      Import of a bank uses the identical syntax of the export command.
      A full example using the normal abbeviations is:

\begin{verbatim}
   im r 5 25 /home/will/downloads/some new bank
\end{verbatim}

      This will look for the directory "some new bank" (spaces are accepted) in
      the download directory of user "will". It will then copy it into bank number
      25 of root number 5. It first checks to ensure that the new named bank
      exists, root 5 exists, and bank 25 is empty.

   \itempar{add root [s]}{cmd!add root}
      Define a new root path, \textsl{s}, and add it to the list of root paths.

   \itempar{remove root [s]}{cmd!remove root}
      De-list the root path named \textsl{s}.

   \itempar{list banks [n]}{cmd!list banks}
      List the instruments and IDs in bank \textsl{n} or the
      current bank/root.

   \itempar{list effects [n]}{cmd!list effects}
      List effect types for \textsl[n].
      If the parameter is the word \textsl{all},
      then list every effect and all its
      presets along with the preset number.

   \itempar{list history [s]}{cmd!list history}
      Displays the recently-used files, including patchsets, scales, and
      states.  If no parameter is given, then this command lists all three
      files in sequence.  The shortest version of this command is
      \texttt{l h p} (for patchsets, which returns the last-seen patchset list).

      \index{at-sign}
      \index{cmd!at-sign}
      \index{cmd!list operator}
      Once that list is displayed,
      the \texttt{@} operator can be used to access
      the item by number.  For example, to load the patch set at location 4 in
      the list:

      \begin{verbatim}
         yoshimi> lo p @4
      \end{verbatim}

   \itempar{list instruments [n]}{cmd!list instruents}
      List all instruments and IDs in bank \textsl{n}
      or the current bank/root.
      Listing instruments will identify the current one with an asterisk, and
      shows the current bank and root one is listing from, and adds a suffix to
      the entry with \textbf{A}, \textbf{S}, or \textbf{P}
      depending on the combination of AddSynth, SubSynth, and PadSynth.

   \itempar{list parts}{cmd!list parts}
      Lists the number of parts available and parts with instruments
      currently installed along with any enabled with the default sound.
      Also shows their audio destination:
      \textsl{M} = main L/R, \textsl{P} = part L/R, \textsl{B} = both, and
      \textsl{-} = disabled or unavailable.
      This way one can tell if an instrument patch is installed even if it is
      not currently usable.
      To avoid unnecessary list length, the default "Simple Sound" is not shown
      unless it is enabled.

   \itempar{list roots}{cmd!list roots}
      Displays all defined root paths.
      Listing roots will identify the current ones with an asterisk.

   \itempar{list setup}{cmd!list setup}
      Displays the current dynamic system settings.

   \itempar{list vector [n]}{cmd!list vector}
      Lists the settings for vector on channel \textsl{n}.

   \itempar{load instrument [s]}{cmd!load instrument}
      Loads an instrument into the current part from the named file.
%     The file-name parameter \textsl{s} is mandatory.

   \itempar{load patchset [s]}{cmd!load instrument}
      Load a complete patch set from a named file, \textsl{s}.
      A variation on this command is \texttt{load patchset @4}, which
      loads the patchset at location 4, the 4th item in the list.

   \itempar{load vector [s]}{cmd!load vector}
      Loads an vector setup from the named file.
      The file-name parameter \textsl{s} is mandatory.

   \itempar{save patchset [s]}{cmd!save patchset}
      Saves the current patchset to the file named \textsl{s}.

   \itempar{save instrument [s]}{cmd!save instrument}
      Saves the instrument of the current part to the named file.
      The file-name parameter \textsl{s} is mandatory.

   \itempar{save setup}{cmd!save setup}
      Save the current dynamic system settings.
      These settings get saved to the state file (we think).

   \itempar{save vector [s]}{cmd!save vector}
      Saves the vector setup to the named file.
      The file-name parameter \textsl{s} is mandatory.

   \itempar{set activate [n]}{cmd!set activate}
      Set part-activate on MIDI program change.
      \textsl{n}=0 disables this feature, and
      1 or any non-zero value enables this feature.
      This feature applies to command line program change as well.

   \itempar{set alsa audio [s]}{cmd!set alsa audio}
      Sets the name of the audio hardware device ALSA looks for.
      Requires a restart of \textsl{Yoshimi}.

   \itempar{set alsa midi [s]}{cmd!set alsa midi}
      Sets the name of the MIDI source ALSA looks for.
      Requires a restart of \textsl{Yoshimi}.

   \itempar{set available [n]}{cmd!set available}
      Set the number of available parts (16, 32, 64).
      Note that 32 and 64 are supported in the newest versions of
      \textsl{Yoshimi}.  Also note that a single two-part vector setup (the
      \textbf{X} vector) requires 32 parts, while the dual two-part vector
      setup (both \textbf{X} and \textbf{Y}) requires 64 parts.

   \itempar{set bank [n]}{cmd!set bank}
      Set current bank to ID \textsl{n}.

   \itempar{set ccbank [n]}{cmd!set ccbank}
      Set the MIDI CC for bank changes (anything other than 0 or 32
      disables MIDI CC).

   \itempar{set ccroot [n]}{cmd!set ccroot}
      Set the MIDI CC for root path changes (values above 119 disable this
      feature).

   \itempar{set extend [n]}{cmd!set extend}
      Set CC value for extended program change (values above 119 disables this
      feature).

   \itempar{set insert effects [n]}{cmd!set insert effects}
      Set insertion effects for editing.
      What are the possible values of \textsl{n}?

   \itempar{set jack midi [s]}{cmd!set jack midi}
      Sets the name of the JACK MIDI source for Yoshimi.
      Requires a restart of \textsl{Yoshimi}.

   \itempar{set jack server [s]}{cmd!set jack server}
      Sets the name of the JACK server Yoshimi tries to connect to.
      Requires a restart of \textsl{Yoshimi}.

   \itempar{set part [n1] program [n2]}{cmd!set part program}
      Load instrument \textsl{n2} into part \textsl{n1}.
      Example: \texttt{set part 4 program 130}

   \itempar{set part [n1] channel [n2]}{cmd!set part channel}
      Set the MIDI channel \textsl{n2} for part \textsl{n1}.
      If the channel number is greater than 15, no further MIDI
      messages will be accepted by that part.

   \itempar{set part [n1] destination [n2]}{cmd!set part destination}
      Set the audio destination of part \textsl{n1}
      to main (\textsl{1}), part (\textsl{2}), both (\textsl{3}).
      Also enables the part if not already enabled.

   \itempar{set preferred audio [s]}{cmd!set preferred audio}
      Set the audio connection type.
      The parameter should be either "jack" or "alsa".

   \itempar{set preferred midi [s]}{cmd!set preferred midi}
      Set the MIDI connection type.
      The parameter should be either "jack" or "alsa".

   \itempar{set - preset [n]}
      Set effect preset.
      Set numbered effect preset.

   \itempar{set program [n]}{cmd!set program}
      Set MIDI program change (0 disables, anything else enables).

   \itempar{set reports [n]}{cmd!set reports}
      Sets the report destination or where messages are displayed, and, to some
      extent, which messages are displayed.  Here are the variations on this
      command that are supported:

      \begin{itemize}
         \item \texttt{set reports gui} or \texttt{s r g}.
            All reports are sent to the GUI console window.
         \item \texttt{set reports stderr} or \texttt{s r s}.
            All reports are sent to stderr.
         \item \texttt{set reports show} or \texttt{s r sh}.
            All messages are displayed.
         \item \texttt{set reports hide} or \texttt{s r h}.
            Non fatal low-level messages are discarded.
         \item \texttt{set reports (any other word or nothing at all)} or
            \texttt{s r (other)}.
            This sets the default condition of sending reports to the CLI and
            displaying all of them.
      \end{itemize}

   \itempar{set root [n]}{cmd!set root}
      Set current root path to ID \textsl{n}.

   \itempar{set shift [n]}{cmd!set shift}
      Set the master key shift for following notes in semitones (+-
      octave, 64 for no shift).

   \itempar{set system effects [n]}{cmd!set system effects}
      Set System Effects for editing.

   \itempar{set vector [n1] x/y cc [n2]}{cmd!set vector cc}
      CC \textsl{n2} is used for channel \textsl{n1} X or Y axis sweep.
      For X, this also enables vector control for the channel.

      \index{vector!features}
      The individual features are now numbered 1-4 and can be
      \index{vector!enable} \texttt{enabled} or
      \index{vector!reverse} \texttt{reversed} (any
      other word disables the feature).
      "Reversed" means that, instead of the X left rising in
      value with increasing CC value, it decreases.
      X right does the opposite of course.

      Feature 1 is always fixed as 7 (volume) and is not reversible.
      Features 2 to 4 can also have the outgoing CC changed to any valid one.
      The vector is just about the only command-line
      entry that starts from 1.

      The original system where bits were ORred together was done to make NRPN
      control as efficient as possible. That hasn't changed, but log messages
      refer to the command-line numbering.

      A more detailed discussion of command-line vector control is presented in
      \sectionref{subsection:vector_command_line}.

   \itempar{set vector [n1] x/y features [n2]}{cmd!set vector features}
      Sets channel \textsl{n1} X or Y features to \textsl{n2}.

   \itempar{set vector [n1] x/y program [l/r] [n2]}{cmd!set vector program}
      Loads program \textsl{n2} to channel \textsl{n1} X or Y
      \textsl{left} or \textsl{right} part.

   \itempar{set vector [n1] x/y control [n2] [n3]}{cmd!set vector control}
      Sets \texttt{n3} CC to use for X or Y feature \texttt{n2} (2, 4, 8).
      \textsl{n3} is the CC to be used for feature number \textsl{n2} on X
      vector channel \textsl{n1}. The \textsl{x} is a sort of hidden parameter
      as the code uses an offset dependent on whether it is \textsl{x} or
      \textsl{y}. Also \textsl{n1} can be omitted in which case it will use the
      last defined channel number. Using alternate words and numbers gives a
      great deal of flexibility like this.

   \itempar{set vector [n] [off]}{cmd!set vector}
      Disables vector control for channel \textsl{n}.

   \itempar{set volume [n]}{cmd!set volume}
      Set the master volume.

   \itempar{reset}{cmd!reset}
      Return to the start-up conditions, if 'y' selected.

   \itempar{stop}{cmd!stop}
      Cease all sound immediately!

   \itempar{? or help}{cmd!help}
      List commands for current mode.  All of the minimum command-line
      abbrevations are capitalised in the help listing.

   \itempar{exit}{cmd!exit}
         Tidy up and close Yoshimi down.

\subsection{Direct Access}
\label{subsec:command_line_direct_access}

   \index{Direct Access}
   \index{cmd!direct access}
   \textsl{Direct Access} is a very low-level access method for most of the
   controls in \textsl{Yoshimi} control.
   It is a test feature accessible only from the command line.
   There are virtually no error checks on direct-access, so one can easily
   crash Yoshimi with out of range values.
   It mostly updates the GUI too.
   Refer to the \texttt{Yoshimi Control Numbers.ods} file that ships with the
   source-code
   for control numbers, actions, and \texttt{GUI\_Interpretations.txt}
   further explanations.
   Also, see \texttt{Effect\_Inserts.txt} for the actual effect controllers and
   \texttt{Envelope\_Inserts} for the enveople controller ranges.

   Sections currently supported by direct access:

   \begin{itemize}
      \item Top level controls
      \item MIDI-learn, used for all operations
      \item Main part controls, now used by GUI for writes
      \item Controllers Window
      \item MIDI CCs window, used for all except incoming real MIDI
      \item Kit edit window
      \item AddSynth Editor
      \item AddSynth Voice Editor
      \item SubSynth Editor
      \item PadSynth Editor
      \item Oscillator
      \item Resonance
      \item Effects
      \item LFOs
      \item Filters
      \item Envelopes
   \end{itemize}

   The remaining sections display as if written to, but don't change anything.

   The official direct access numbering system is zero based (and will remain
   so).

%  This matches the results from using the middle mouse button in the GUI
%  (assuming it was enabled at build time). Commands that are purely within
%  the GUI, e.g window changes are recognised but not acted on.
%
%  An alternative way to find out the correct command sequence is to click on
%  a control with any button *except* the middle one. Trailing '255's can be
%  ommitted if there are no following parameters set.
%
%  In some situations par2 is used as a miscMSG ID for passing text values
%  between threads but synchronised via the RT thread.

   The format of the direct-access command is:

   \begin{verbatim}
      direct [value] [type] [control] [part] [kit] [engine] [insert] [parameter] [par2]
   \end{verbatim}

   \begin{itemize}
      \item \texttt{direct} is the actual command name, must be typed in full.
      \item \texttt{value} is the value that will be set, may be negative and
         floating point (looks for the devimal point).
      \item \texttt{type} is a subset of the command type: 0 for read and 64 for
         write. Sending 1 will just echo the bytes you sent.  Sending 3 will
         turn it into a MIDI-learn command.
      \item \texttt{control} is the number representing the control being
         changed or queried.
      \item \texttt{part} is the number of the part being changed.
   \end{itemize}

   All the parameters above are mandatory. The following must also be entered in
   sequence when needed.

   \begin{itemize}
      \item \texttt{kit} is a part's kit number (if enabled otherwise zero).
         also used for effects.
      \item \texttt{engine} is the synth engine being adjusted.
      \item \texttt{insert} is harmonics, LFOs, filters etc.
      \item \texttt{parameter} is subtype associated with some inserts.
      \item \texttt{par2} is an offset for some parameters.
   \end{itemize}

   A variation of the above is when one wishes to find the limits of controls.
   While it is possible to do this with the above method, it is complex and
   error-prone so the following options have been implemented:

   \begin{verbatim}
      direct limits min
      direct limits max
      direct limits def
   \end{verbatim}

   These are followed by the usual (control) (part) etc.  Most of these are independently declared. The remainder have the values: min = 0, max = 127, and def = 0.

   CLI direct access to show limits returns a single value dependent on the
   initial text and parameters.  Some examples:
   The control number for volume, part number, will return:

   \begin{verbatim}
      direct limit max 0 0
      Max 127.000000
   \end{verbatim}

For panning:

   \begin{verbatim}
      direct limit  def 1 0
      Default 64.000000
   \end{verbatim}

For keyshift, top level controls:

   \begin{verbatim}
      direct limit min 35 240
      min -36.000000
   \end{verbatim}

From the command line you can now read the VU levels:

   \begin{verbatim}
      direct 0 0 200 240 {part number}
   \end{verbatim}

will show the numbered part's peak level.
It will be negative if the part is not enabled.

   \begin{verbatim}
      direct 0 0 201 240 0
   \end{verbatim}

gives the main output LH peak.

   \begin{verbatim}
      direct 0 0 201 240 1
   \end{verbatim}

gives the main output RH peak.

   \begin{verbatim}
      direct 0 0 202 240 0
   \end{verbatim}

gives the main output LH RMS.

   \begin{verbatim}
      direct 0 0 202 240 1
   \end{verbatim}

gives the main output RH RMS.

   Obviously, these are fluctuating values, but give an idea of what is
   happening.  Those zeros at the start are necessary; and the commands are
   exactly what the GUI requests when updating the displays.

%-------------------------------------------------------------------------------
% vim: ts=3 sw=3 et ft=tex
%-------------------------------------------------------------------------------
