%-------------------------------------------------------------------------------
% yum_bank_collection
%-------------------------------------------------------------------------------
%
% \file        yum_bank_collection.tex
% \library     Documents
% \author      Chris Ahlstrom
% \date        2015-06-20
% \update      2015-07-18
% \version     $Revision$
% \license     $XPC_GPL_LICENSE$
%
%     Provides the banks_collection section of yoshimi-user-manual.tex.
%
%-------------------------------------------------------------------------------

\section{Banks Collection}
\label{sec:banks_collection}

   In this section, we attempt to collect and summarise all of the existing
   banks for \textsl{Yoshimi} and \textsl{ZynAddSubFX}
   that we can find.  Many of them
   are supplied by the two projects.

   Between all of the collections, there is a large amount of duplication.
   There is also semi-duplication, with slight variations on the same basic
   instrument.
   Various Linux distributions which package \textsl{ZynAddSubFX}
   and \textsl{Yoshimi}
   might add some banks to their versions of these packages.
   Thus, there are far more sound settings than we can discuss and categorise.

   One thing we're looking for is a good General MIDI (GM) bank for
   \textsl{Yoshimi}.  As part of our \textsl{Yoshimi Cookbook}
   \cite{cookbook}, we include a basic General MIDI bank for.
   However, there are number of patches with no good implementation in it.

\subsection{Yoshimi Banks}
\label{subsec:banks_collection_yoshimi}

   \textsl{Yoshimi} comes with the following banks, which may be found in
   \texttt{/usr/share/yoshimi/banks} as installed by the installer.  In this
   case, it is the Debian installer.

   \begin{enumber}
      \item \textbf{Arpeggios}.
         Also in ZynAddSubFX.
      \item \textbf{Bass}.
         Also in ZynAddSubFX.
      \item \textbf{Brass}.
         Also in ZynAddSubFX.
      \item \textbf{chip}.
      \item \textbf{Choir\_and\_Voice}
         Also in ZynAddSubFX, slightly different bank name.
      \item \textbf{Drums}.
         Also in ZynAddSubFX, but with only one drum kit included.
      \item \textbf{Dual}.
         Also in ZynAddSubFX.
      \item \textbf{Fantasy}.
         Also in ZynAddSubFX.
      \item \textbf{Guitar}.
         Also in ZynAddSubFX.
      \item \textbf{Misc}.
         Also in ZynAddSubFX.
      \item \textbf{Noises}.
         Also in ZynAddSubFX.
      \item \textbf{Organ}.
         Also in ZynAddSubFX.
      \item \textbf{Pads}.
         Also in ZynAddSubFX.
      \item \textbf{Plucked}.
         Also in ZynAddSubFX.
      \item \textbf{Reed\_and\_Wind}.
         Also in ZynAddSubFX, slightly different bank name.
      \item \textbf{Rhodes}.
         Also in ZynAddSubFX.
      \item \textbf{Splited}.
         Also in ZynAddSubFX, slightly different bank name.
      \item \textbf{Strings}.
         Also in ZynAddSubFX.
      \item \textbf{Synth}.
         Also in ZynAddSubFX.
      \item \textbf{SynthPiano}.
         Also in ZynAddSubFX.
      \item \textbf{The\_Mysterious\_Bank}.
         Also in ZynAddSubFX, slightly different bank name.
         ZynAddSubFx has three more mysterious banks (see next section).
      \item \textbf{Will\_Godfrey\_Collection}.
      \item \textbf{Will\_Godfrey\_Companion}.
   \end{enumber}

\subsection{Additional ZynAddSubFX Banks}
\label{subsec:banks_collection_zynaddsubfx}

   \textsl{ZynAddSubFX} comes with the following banks, which may be found in
   the source code \cite{zynsource} or installation packages of this project.
   \textsl{ZynAddSubFX} has some of the same
   banks (as far as we can tell) as \textsl{Yoshimi}, but with the following
   additions:

   \begin{enumber}
      \item \textbf{Companion}.
      \item \textbf{Cormi\_Noise and Cormi\_Sound} \cite{cormi}.
      \item \textbf{Laba170bank}.
      \item \textbf{olivers-100}.
         Some very good instruments, including sitar
         and steel drums.
      \item \textbf{the\_mysterious\_bank}.
      \item \textbf{the\_mysterious\_bank\_2}.
      \item \textbf{the\_mysterious\_bank\_3}.
      \item \textbf{the\_mysterious\_bank\_4}.
   \end{enumber}

\subsection{Additional Banks}
\label{subsec:banks_collection_additional}

   Here are some additional banks we have found, or have built ourselves.
   It often happens that, later on, a site is no longer available.
   Or we forget from whence we got the banks.
   In these cases, the banks are stored in the \texttt{contrib/banks}
   directory of this project.

   \begin{enumber}
      \item \textbf{Alex\_J}
         The site seems to be gone/expired. So one will find these in the
         "contrib/banks" directory for safekeeping.
      \item \textbf{Bells}
         *
      \item \textbf{C\_Ahlstrom}
         These are mine, but not yet made into a systematic bank.
         They are included with this document.
      \item \textbf{Chromatic Percussion}
         *
      \item \textbf{Drums\_DS}
         *
      \item \textbf{Electric Piano}
         *
      \item \textbf{Flute}
         *
      \item \textbf{folderol collection}
         \cite{folderol}, also found at \cite{zyndemos}.
      \item \textbf{Internet Collection}
         *
      \item \textbf{Leads}
         *
      \item \textbf{Louigi\_Verona\_Workshop}
         The site seems to be gone/expired. So one will find these in the
         "contrib/banks" directory for safekeeping.
      \item \textbf{Misc Keys}
         *
      \item \textbf{mmxgn Collection} \cite{mmxgn}
      \item \textbf{Piano}
         *
      \item \textbf{RB Zyn Presets}
         *
      \item \textbf{Vanilla}
         See \cite{zyndemos} for this bank, and for some demonstration files
         of \textsl{ZynAddSubFX} sounds, and some other nice links.
      \item \textbf{VDX}
         *
      \item \textbf{x31eq.com} \cite{x31eq}
      \item \textbf{XAdriano Petrosillo}
         *
      \item \textbf{Zen Collection}
         *
   \end{enumber}
   * The source of these banks is currently unknown.


   To find out how to integrate these into one's installation, see
   \sectionref{subsec:banks_and_roots_roots}, and
   \sectionref{subsubsec:menu_patch_sets_show_patch_banks}

%-------------------------------------------------------------------------------
% vim: ts=3 sw=3 et ft=tex
%-------------------------------------------------------------------------------
